---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(stringr)
library(ggvis)

```

# Analysis of a sample of Monster.com job postings

Sample of a kaggle dataset of job postings. 


``` {r read, echo=TRUE} 

# Load data and drop unnecessary columns

df <- read.csv('monster_com-job_sample.csv', stringsAsFactors = FALSE, na.strings = '') 

df <- df %>% select(job_description,
                    job_type,
                    location,
                    organization,
                    salary,
                    sector)

glimpse(df)

```


Some rows have the location in the organization column, so we have to move those over. To do this, we'll get rid of the zip codes and then move the 'organization' to the 'location' column if it can be found anywhere in the location column.


```{r cars, echo=TRUE}

# how many in organization is found in location before removing zip code

sum(df$organization %in% df$location & !is.na(df$organization))

# remove zip codes

df$location <- sub('\\d{5}', '', df$location)
df$organization <- sub('\\d{5}', '', df$organization)

# how many are found after

sum(df$organization %in% df$location & !is.na(df$organization))

# move the values over

df[((df$organization %in% df$location) && (!is.na(df$organization))), 'location'] <-
  df[((df$organization %in% df$location) && (!is.na(df$organization))), 'organization']

# separate state into its own column

df$state <- lapply(1:length(df$location), function(x) {
  if (str_extract(df[x, 'location'], '[[:upper:]]{2}') %>% str_trim() %in% state.abb) {
  str_extract(df[x, 'location'], '[[:upper:]]{2}') %>% str_trim()
  }
  else
  NA
}) %>% unlist()

# count of jobs grouped by state
```

``` {r state, echo=TRUE, fig.width = 10}
df %>% group_by(state) %>% summarize(jobs = n()) %>% ggvis(~state, ~jobs, fill := 'coral') %>% layer_bars()
```

The dataset is 22,000 observations, but 84.3% of those don't have salary information. Those with salary information have it in different formats. Some are hourly, some are yearly, some have a range of salaries, and some don't specify an amount. I'll have to try to separate ranged salaries and standardize hourly/yearly salaries.

```{r pressure, echo=TRUE, message=TRUE, warning=FALSE}

# remove commas from salary column

df$salary <- gsub(',', '', df$salary)

# add salary type of yearly or hourly

df$salary_type <- lapply(1:length(df$salary), function(x) {
  if (grepl('year', df[x, 'salary']) == TRUE) {
    'yearly'
  }
  else if (grepl('hour', df[x, 'salary']) == TRUE)
    'hourly'
  else
    NA
}) %>% unlist()

# add min column

df$min <- lapply(1:length(df$salary), function(x) {
  if (length(str_extract_all(df[x, 'salary'], '\\d+.\\d+')[[1]]) > 1  && (!is.na(df[x, 'salary']))) {
    str_extract_all(df[x, 'salary'], '\\d+.\\d+')[[1]][1]
  }
  else
    NA
}) %>% unlist() %>% as.numeric()

df$max <- lapply(1:length(df$salary), function(x) {
  if (length(str_extract_all(df[x, 'salary'], '\\d+.\\d+')[[1]]) > 1  && (!is.na(df[x, 'salary']))) {
    str_extract_all(df[x, 'salary'], '\\d+.\\d+')[[1]][2]
  }
  else
    NA
}) %>% unlist() %>% as.numeric()

# some of the mins are 0, remove those

df[df$min == 0 & !is.na(df$min), 'min'] <- NA

# convert yearly salaries to hourly

for (i in 1:nrow(df)) {
  if (df[i, 'salary_type'] == 'hourly' && !is.na(df[i, 'salary_type'])) {
    df[i, 'max'] <- df[i, 'max'] * 2080
    df[i, 'min'] <- df[i, 'min'] * 2080
  }
}

# new column with mean of min/max

df <- df %>% mutate(avg = (min + max) / 2)

# find outliers in wages

df %>% arrange(desc(avg)) %>% head(n=10) %>% select(avg)

# remove any average wage over 1,000,000 and under 10,000. They were probably entered incorrectly

df$avg <- lapply(1:length(df$avg), function(x) {
  if (df[x,'avg'] >= 1000000  && (!is.na(df[x, 'avg']))) {
    NA
  } else if (df[x,'avg'] <= 10000  && (!is.na(df[x, 'avg']))) {
    NA
  }
  else
    df[x,'avg']
}) %>% unlist() %>% as.numeric()

```

```{r wages, echo=T, fig.width=10}
# average of wages by state (with over 5 salaries)

df %>% group_by(state) %>% filter(sum(!is.na(avg)) >5) %>% summarize(wages = mean(avg, na.rm = T)) %>% ggvis(~state, ~wages, fill := 'coral') %>% layer_bars()


```

```{r sector, echo=T, fig.width=10, fig.height=8}
# average of wages by sector (with over 5 salaries)

df %>% group_by(sector) %>% filter(!is.na(sector), !is.na(avg), sum(!is.na(avg)) > 5) %>% summarize(wages = mean(avg, na.rm = T)) %>% ggvis(~wages, ~sector, fill := 'coral') %>% layer_rects(x2=0,height = band())

```

